---
title: "RinActionChap07 Basic statics"
author: "Tony"
date: "2016年1月2日"
output: html_document
---
### Descriptive statistics

####Amenagerie of methods
```{r}
vars <- c("mpg", "hp", "wt")
head(mtcars[vars])
summary(mtcars[vars])

mystats <- function(x, na.omit=FALSE) {
  if(na.omit)
    x <- x[!is.na(x)]
  m <- mean(x)
  n <- length(x)
  s <- sd(x)
  skew <- sum((x-m)^3/s^3)/n
  kurt <- sum((x-m)^4/s^4)/n - 3
  return(c(n=n, mean=m, stdev=s, skew=skew, kurtosis=kurt))
}

sapply(mtcars[vars], mystats)
```

hmisc pastecs psych

```{r}
library(Hmisc)
describe(mtcars[vars])
library(pastecs)
stat.desc(mtcars[vars])
library(psych)
describe(mtcars[vars])
Hmisc::describe(mtcars[vars])
```

####Descriptive statistics by group

```{r}
# aggregate(mtcars[vars], by=list(am=mtcars$am), mean)
# dstats <- function(x)(c(mean=mean(x), sd=sd(x)))
# by(mtcars[vars], mtcars$am, dstats)
# library(doBy)
# summaryBy(mpg+hp+wt~am, data=mtcars, FUN=mystats)
# library(psych)
# describe.by(mtcars[vars], mtcars$am)
# library{reshape}
# dstats <- function(x)(c(n=length(x), mean=mean(x), sd=sd(x)))
# dfm <- melt(mtcars, measure.vars=c("mpg", "hp", "wt"),
#             id.vars=c("am", "cyl"))
# cast(dfm, am + cyl + variable ~ ., dstats)
```

###Frequency and contingency tables
```{r}
library(vcd)
head(Arthritis)

```

Function | Description
-------- | -----------
table(var1, var2, …, varN)  | Creates an N-way contingency table from N categorical variables (factors)
xtabs(formula, data) | Creates an N-way contingency table based on a formula and a matrix or data frame
prop.table(table, margins) | Expresses table entries as fractions of the marginal table defined by the margins
margin.table(table, margins) | Computes the sum of table entries for a marginal table defined by the margins
addmargins(table, margins) | Puts summary margins (sums by default) on a table
ftable(table) | Creates a compact "flat" contingency table

```{r}
mytable <- with(Arthritis, table(Improved))
mytable
prop.table(mytable)
prop.table(mytable)*100

mytable <- xtabs(~ Treatment+Improved, data=Arthritis) 
mytable
margin.table(mytable, 1)
prop.table(mytable, 1)
margin.table(mytable, 2)
prop.table(mytable, 2)
addmargins(mytable)
library(gmodels)
CrossTable(Arthritis$Treatment, Arthritis$Improved)
```
####Tests of independence

```{r}
library(vcd)
mytable <- xtabs(~Treatment+Improved, data=Arthritis)
chisq.test(mytable)
fisher.test(mytable)
mytable <- xtabs(~Improved+Sex, data=Arthritis)
chisq.test(mytable)
mytable <- xtabs(~Treatment+Improved+Sex, data=Arthritis)
mantelhaen.test(mytable)
           
```

####Measures of association

```{r}
library(vcd)
mytable <- xtabs(~Treatment+Improved, data=Arthritis)
assocstats(mytable)
```

####Converting tables to flat files

```{r}
table2flat <- function(mytable) {
  df <- as.data.frame(mytable)
  rows  <- dim(df)
  cols  <- dim(df)[2]
  x <- NULL
  for (i in 1:rows){
    for (j in 1:df$Freq[i]){
      row <- df[i,c(1:(cols-1))]
      x <- rbind(x,row)
      }
    }
  row.names(x)<-c(1:dim(x)[1])
  return(x)
}

treatment <- rep(c("Placebo", "Treated"), times=3)
improved <- rep(c("None", "Some", "Marked"), each=2)
Freq <- c(29,13,7,17,7,21)
mytable <- as.data.frame(cbind(treatment, improved, Freq))
mydata <- table2flat(mytable)
head(mydata)
```

###Correlations 
####Types of correlations
- pearson
- spearman
- kendall
- `cor(x, user=, method=)`
- partial correlations `pcor(u, S)`

```{r}
states <- state.x77[,1:6]
cov(states)
cor(states)
cor(states, method = "spearman")
x <- states[,c("Population", "Income", "Illiteracy", "HS Grad")]
y <- states[,c("Life Exp", "Murder")]
cor(x, y)
```

####Testing correlations for significance

```{r}
cor.test(states[,3], states[,5])
library(psych) 
corr.test(states, use="complete")

```
####Visualizing correlations
scatter plots

###t-test
`t.test(y ~ x, data)`

```{r}
library(MASS)
t.test(Prob ~ So, data=UScrime)
sapply(UScrime[c("U1","U2")], function(x)(c(mean=mean(x),sd=sd(x))))
with(UScrime, t.test(U1, U2, paired=TRUE))
```

###Nonprarametric tests of group differences
- `wilcox.test(y ~ x, data)`
- `kruskal.test(y ~ A, data)`
- `friedman.test(y ~ A | B, data)`