---
title: "RinActionChap02 Creating a dataset"
author: "Tony"
date: "2016年1月1日"
output: html_document
---

##Basic 
###Data structure
####Vectors
```{r}
a <- c(1, 2, 5, 3, 6)
b <- c("one", "two")
c <- c(TRUE, FALSE)
a[3]
a[1:3]
```

####Matrices
```{r}
y <- matrix(1:20, nrow = 5, ncol = 4)
y
cells <- c(1, 26, 24, 68)
rnames <- c("r1", "r2")
cnames <- c("c1", "c2")
mymatrix <- matrix(cells, nrow = 2, ncol = 2, byrow = TRUE, dimnames = list(rnames, cnames))
mymatrix

mymatrix <- matrix(cells, nrow = 2, ncol = 2, byrow = FALSE, dimnames = list(rnames, cnames))
mymatrix

x <- matrix(1:10, nrow = 2)
x
x[2,]
x[,2]
x[1,3]
x[1, c(4, 5)]


```

####Arrays
Arrays are similar to matrices but can have more than two dimensions

```{r}
dim1 <- c("A1", "A2")
dim2 <- c("B1", "B2", "B3")
dim3 <- c("C1", "C2", "C3", "C4")
z <- array(1:24, c(2, 3, 4), dimnames=list(dim1, dim2, dim3))
z
```

####Data frames
different columns can contain different modes of data

```{r}
patientID <- c(1, 2, 5, 4)
age <- c(25, 34, 28, 52)
diabetes <- c("Type1", "Type2", "Type1", "Type1")
status <- c("Poor", "Improved", "Excellent", "Poor")
patientdata <- data.frame(patientID, age, diabetes, status)
patientdata
patientdata[1 : 2]
patientdata[c("diabetes", "status")]
patientdata$age
table(patientdata$diabetes, patientdata$status)
patientdata <- data.frame(patientID, age, diabetes, status, row.names=patientID)
patientdata
```

```{r}
attach(mtcars)
  summary(mtcars$mpg)
  plot(mtcars$mpg, mtcars$disp)
detach(mtcars)
```

with(): assignment will only exist within function brackets unless use `<<-` 

```{r}
with(mtcars, {
  stats <- summary(mpg, disp, wt)
  stats2 <<- summary(mpg, disp, wt)
})
stats2
```

####Factors
```{r}
patientID <-c(1, 2, 3, 4)
age <- c(25, 34, 28, 52)
diabetes <- c("Type1", "Type2", "Type1", "Type1")
status <- c("Poor", "Improved", "Excellent", "Poor")
diabetes <- factor(diabetes)
status <- factor(status, order=TRUE)
patientdata <- data.frame(patientID, age, diabetes, status) 
str(patientdata)
summary(patientdata)
```

####Lists
A list is an ordered collection of objects

```{r}
g <- "hello"
h <- c(25,12)
j <- matrix(1:10, nrow = 2)
k <- c("one", "two")
mylist <- list(title=g, ages = h,j,k)
mylist
```

###Data input
####Entering data from the keyboard
```{r}
mydata <- data.frame(age=numeric(0),
                     gender=character(0), weight=numeric(0))
mydata <- edit(mydata)

```

####Import data from a delimited text file
```
mydataframe <- read.table(file, header = logical_value,
                          seq="delimiter", row.names = "name");
grades <- read.table("studentgrades.csv", header=TRUE, sep=",",   row.names="STUDENTID")
```

####Import data from Excel

```
install.packages("RODBC")
library(RODBC)
channel <- odbcConnectExcel("myfile.xls")
mydataframe <- sqlFetch(channel, "mysheet")
odbcClose(channel)
```

#### XML
XML package 
http://www.omegahat.org/RSXML/

####Webscraping
readLines()
grep()
gsub()
RCurl and XML package

####SPSS
Hmisc package
spss.get()

####SAS
read.ssd() in foreign package
sas.get() in Hmisc

####Stata
read.dta() in foreign package

####netCDF
```
library(ncdf) 
nc <- nc_open("mynetCDFfile") 
myarray <- get.var.ncdf(nc, myvar)
```

####HDF5
hdf5 package

####DBMSs
RODBC package

####Stat/Transfer
It can transfer data between 34 data formats.

###Annotating datasets
```
names(patientdata)[2] <- "Age at hospitalization (in years)"
patientdata$gender <- factor(patientdata$gender,levels = c(1,2),labels = c("male", "female"))
```

###Functions 

```
length()
dim()
str()
class()
mode()
names()
c()
cbind()
rbind()
object
head()
tail()
ls()
rm()
newobject <- edit(object)
fix(object)   Edit in place
```
