---
title: "RinActionChap04 Basic data management"
author: "Tony"
date: "2016年1月1日"
output: html_document
---
###A working example

```{r}
manager <- c(1, 2, 3, 4, 5)
date <- c("10/24/08", "10/28/08", "10/1/08", "10/12/08", "5/1/09")
country <- c("US", "US", "UK", "UK", "UK")
gender <- c("M", "F", "F", "M", "F")
age <- c(32, 45, 25, 39, 99)
q1 <- c(5, 3, 3, 3, 2)
q2 <- c(4, 5, 5, 3, 2)
q3 <- c(5, 2, 5, 4, 1)
q4 <- c(5, 5, 5, NA, 2)
q5 <- c(5, 5, 2, NA, 1)
leadership <- data.frame(manager, date, country, gender, age, 
                         q1, q2, q3, q4, q5, stringsAsFactors =  FALSE)

```

###Create new variables
- `+`
- `-`
- `*`
- `/`
- `^` or `**`
- `x%%y` mod
- `x%/%y` Integer division

```{r}
mydata <- data.frame(x1 = c(2,2,6,4),
                     x2 = c(3,4,2,8))
mydata$sumx <- mydata$x1 + mydata$x2
mydata$meanx <- (mydata$x1 + mydata$x2)/2

attach(mydata)
mydata$sumx <- x1+x2
mydata$meanx <- (x1+x2)/2
detach(mydata)

mydata <- transform(mydata, 
                    sumx <- x1+x2,
                    meanx <- (x1+x2)/2)
mydata
```

###Recoding variables
`<` `<=` `>` `>=` `==` `!=` `!x` `x | y` `x & y` `isTRUE(x)`

```{r}
leadership$age[leadership$age == 99] <- NA
leadership$agecat[leadership$age >75] <- "Elder"
leadership$agecat[leadership$age >= 55 &&
                    leadership$age <=75] <- "Middle Aged"
leadership$agecat[leadership$age < 55] <- "Young"
# or
leadership <- within(leadership, {
  agecat <- NA
  agecat[age > 75] <- "Elder"
  agecat[age >= 55 & age<=75] <- "Middle Aged"
  agecat[age < 55] <- "Young1"})
# leadership
```

###Renaming variables
- Use `fix(leadership)` to open an editor
```{r}
library(reshape)
leadership <- rename(leadership, 
                     c(manager="managerID"))
# names(leadership)[6:10] <- c("item1", "item2", "item3", "item4", "item5")
# leadership
```

###Missing values
```{r}
x <- c(1,2,3,NA)
# is.na(x)
# is.na(leadership)

y <- sum(x, na.rm = TRUE)

# only complete cases
newdata <- na.omit(leadership)


```

###Date value
- `%d` 01-31
- `%a` Mon
- `%A` Monday
- `%m` 00-12
- `%b` Jan
- `%B` January
- `%y` 07
- `%y` 2007

```{r}
mydates <- as.Date("2007-06-22", "2004-02-13")
strdate <- c("01/05/1965", "08/16/1975")
dates <- as.Date(strdate, "%m/%d/%y")
myformat <- "%m/%d/%y"
leadership$date <- as.Date(leadership$date, myformat)
leadership
today <- Sys.Date()
format(today, format="%B %d %Y")
format(today, format="%A")
startdate <- as.Date("2004-02-13")
enddate   <- as.Date("2011-01-22")
days <- enddate - startdate
str(days)
days
strDates <- as.character(today)
# help("as.Date")
# help("strftime")
#fCalendar package

```

###Type conversions
####Test and convert function
- `is...()`
- `as...()`
- numeric character vector matrix data.frame factor logical

###Sorting data
```{r}
attach(leadership)
newdata <- leadership[order(gender, age),]
newdata <- leadership[order(gender, -age),]
detach(leadership)

```

###Merging datasets
```{r}
# total <- merge(dataframeA, dataframeB, by="ID")
# total <- merge(dataframeA, dataframeB, by=c("ID","Country"))
# total <- cbind(A, B)
# total <- rbind(dataframeA, dataframeB)
```

###Subsetting datasets

####Selcting variables
```{r}
myvars <- c("q1", "q2", "q3", "q4", "q5")
newdata <- leadership[myvars]
```

####Excluding(dropping) variables
```{r}
myvars <- names(leadership) %in% c("q3", "q4")
newdata <- leadership[!myvars]
newdata$q5 <- NULL
# newdata
```

####Selecting observations
```{r}
leadership$date <- as.Date(leadership$date, "%m/%d/%y")
startdate <- as.Date("2009-01-01")
enddate   <- as.Date("2009-10-31")
newdata <- leadership[which(leadership$date >= startdate & leadership$date <= enddate),]
newdata
```

#### subset()
```{r}
newdata = subset(leadership, age >=35 | age < 24, 
                 select=c(q1, q2, q3, q4,q5))

newdata = subset(leadership, gender = "M" & age > 25, 
                 select= gender:q4)
newdata
```

####Random samples
```{r}
mysample <- leadership[sample(1:nrow(leadership)), 3]
mysample
#sampling package
#survey package

```

####Using SQL
```{r}
library(sqldf)
newdf <- sqldf("select * from mtcars where carb=1 order by mpg")
sqldf("select avg(mpg) as avg_mpg, avg(disp) as avg_disp, gear 
      from mtcars where cyl in (4, 6) group by gear")
```
