---
title: "RinActionChap05 Advanced data management"
author: "Tony"
date: "2016年1月1日"
output: html_document
---

### Numerical and character functions
####Mathematical function
- abs(x)
- sqrt(x)
- ceiling(3.5) = 4
- floor(3.5) = 3
- trunc(3.5) = 3
- round(3.456, digits=2) = 3.46
- signif(3.475, digits=2) = 3.5
- cos(x), sin(x), tan(x)
- acos, asin, atan
- cosh, sinh tanh
- acosh, asinh, atanh
- log(x, base=n)
- log(x) natural logarithm
- log10(x)
- exp(x)

####Statistical function
- mean(c(...))
- median(c(...))
- sd(x) standard deviation 标准差：方差的平方根
- var(x) Variance 方差：各值和平均值差值的平方的平均值
- mad(x) Median absolute deviation：先求出中位数，然后各个数和中位数的差绝对值再求中位数
- quantile(x, probs) TODO:具体分析
- range(x)
- sum(x)
- diff(c(1,3,7)) = c(2,5)
- min(x)
- max(x)
- scale(x, center = TRUE. scale= TRUE)

####Probability functions
In R, probality function take the form 
{dpqr}sitribution_abbreviation()

- d = density
- p = distribution function
- q = quantile function
- r = random generation(random deviates)

Distribution | Abbreviation | Distribution | Abbreviation
------------ | ------------ | ------------ | ------------
Beta | beta | Logistic | logis
Binomial | binom | Multinomial | multinom
Cauchy | cauchy | Negative binomial |nbinom
Chi-squared (noncentral) | chisq | Normal |norm
Exponential | exp | Poisson | pois
F | f | Wilcoxon Signed Rank | signrank
Gamma | gamma | T | t
Geometric | geom | Uniform | unif
Hypergeometric | hyper | Weibull | weibull
Lognormal | lnorm | Wilcoxon Rank Sum |wilcox


```{r}
x <- pretty(c(-3, 3), 30)
y <- dnorm(x)
plot(x, y, 
     type = "l",
     xlab = "Normal Deviate", 
     ylab = "Density",
     yaxs = "i"
     )
```


```{r}
runif(5)
# by setting the seed manually, you're able to reproduce your results
set.seed(1234)
runif(5)
set.seed(1234)
runif(5)
```

####Character functions
- nchar(x) count the number of characters of x
- substr(x, start, stop): 
  x <- "abcdef" 
  substr(x, 2, 4) returns "bcd"
  substr(x, 2, 4) <- "2222", now x is "a222ef"
- grep(pattern, x, ignore.case=FALSE, fixed=FALSE) fixed=FALSE pattern is a regular expression
- sub(pattern, x, replacement, x, ignore.case=FALSE, fixed=FALSE)
- strsplit(x, split, fixed=FALSE)
- paste(... sep="")
  paste("x", 1:3, sep="M") returns c("xM1", "xM2", "xM3")
- toupper(x)
- tolower(x)

####Other userful functions
```{r}
a <- 1:3
length(a)
seq(1,10,2)
rep(1:3, 2)
cut(rep(1:3, 2), 2)
pretty(rep(1:10,2), 3)
#cat(..., file = "myfile", append = FALSE)
```

###A solution for out data management challenge
```{r}
options(digits=2)
Student <- c("John Davis", "Angela Williams", "Bullwinkle Moose", 
             "David Jones", "Janice Markhammer", "Cheryl Cushing",
             "Reuven Ytzrhak", "Greg Knox", "Joel England",
             "Mary Rayburn")
Math <- c(502, 600, 412, 358, 495, 512, 410, 625, 573, 522)
Science <- c(95, 99, 80, 82, 75, 85, 80, 95, 89, 86)
English <- c(25, 22, 18, 15, 20, 28, 15, 30, 27, 18)
roster <- data.frame(Student, Math, Science, English,
                     stringsAsFactors=FALSE)
z <- scale(roster[,2:4])
score <- apply(z, 1, mean)
roster <- cbind(roster, score)
y <- quantile(score, c(.8, .6, .4, .2))
roster$grade[score >= y[1]] <- "A"
roster$grade[score < y[1] & score >= y[2]] <- "B" 
roster$grade[score < y[2] & score >= y[3]] <- "C" 
roster$grade[score < y[3] & score >= y[4]] <- "D" 
roster$grade[score < y[4]] <- "F"
name <- strsplit(roster$Student, " ")
lastname <- sapply(name, "[", 2)
firstname <- sapply(name, "[", 1)
roster <- cbind(firstname, lastname, roster[,-1])
roster
```

###Control flow
```{r}
for (i in 1:3) print("Hello")
i <- 3
while (i >0) 
{
  print("Hello")
  i <- i-1
}

ifelse (TRUE, print("1"), print("0"))

feelings <-c("sad", "afraid")
for (i in feelings)
  print(
    switch(i, 
           happy = "h",
           afraid = "A",
           sad = "s"
           )
  )
```

####User-written functions
```{r}
mystats <- function(x, parametric=TRUE, print=FALSE)
{
  if (parametric)
  {
    center <- mean(x);
    spread <- sd(x)
  }
  else 
  {
    center <- median(x);
    spread <- mad(x)
  }
  
  if (print & parametric)
  {
    cat("Mean=", center, "\n", "SD=", spread, "\n" )
  }
  else if (print & !parametric)
  {
    
    cat("Median=", center, "\n", "MAD=", spread, "\n" )
  }
  result <- list(center=center, spread=spread)
  return(result)
}

set.seed(1234)
x <- rnorm(500)
y <- mystats(x)
y
```

####Aggregation and restructuring
```{r}
cars <- mtcars[1:5, 1:4]
t(cars)

options(digits = 3)
attach(mtcars)
aggdata <- aggregate(mtcars, by=list(cyl, gear), FUN=mean, na.rm=TRUE)
aggdata

# reshape package
library(reshape)
# md <- melt(mydata, id=(c("id", "time")))
# newdata <- cast(md, fomula, FUN)
```